---
title: "Reference dataframes for the acled report"
date: '2022-06-12'
output: 
  html_document:
    code_download: true
    theme: readable
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: false
    collapsed: false
    code_folding: hide
always_allow_html: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
library(tidyverse)
library(janitor)
library(lubridate)
library(stringi)
library(readxl)
library(sf)
library(tidytext)
```



## Townships

From 2021 version of the 5Ws. Administrative level 3 place codes. Public. [Link](https://raw.githubusercontent.com/food-security-cluster-myanmar/exploratory-data-analysis-acled-fsc/master/data/townships.csv). 

```{r}
# Townships, from geoadmins 
townships <- read_excel("FSC 5W 2021 - GEOADMINS_final 19.xlsx",
           sheet = "GEOADMINS") %>% 
  clean_names() %>% 
  select(admin1pcode_4:admin3pcode) %>% 
  rename(admin1_pcode = admin1pcode_4,
         admin3_pcode = admin3pcode,
         state_name   = state_5,
         township_name  = county) %>% 
  remove_empty() %>% 
  mutate(admin3_pcode = ifelse(state_name == "Magway" & township_name == "Minhla", "MMR009013", admin3_pcode),
         admin3_pcode = ifelse(state_name == "Yangon" & township_name == "Htantabin", "MMR013006", admin3_pcode))

townships %>% 
  write_csv("./data/townships.csv")

```

## Shapefiles 

MIMU shapefiles. Administrative levels 1 and 3 shapefiles. Public. Links for [Admin level 1](https://geonode.themimu.info/layers/geonode%3Ammr_polbnda2_adm1_mimu_250k) [Admin level 3](https://geonode.themimu.info/layers/geonode%3Ammr_polbnda_adm3_mimu_250k).  

```{r}
# shapefiles
pcode3_shape <- st_read("./mmr_polbnda_adm3_mimu_250k/mmr_polbnda_adm3_mimu_250k.shp", quiet = TRUE) %>% 
 rename(state = ST, 
        admin1_pcode = ST_PCODE,
        township = TS,
        admin3_pcode = TS_PCODE) %>% 
 mutate(admin3_pcode = ifelse(str_detect(township, "Hlaingtharya"), "MMR013008", admin3_pcode))

pcode1_shape <- st_read("./mmr_polbnda2_adm1_mimu_250k/mmr_polbnda2_adm1_mimu_250k.shp", quiet = TRUE) %>% 
 rename(state = ST, 
        admin1_pcode = ST_PCODE) %>% st_as_sf()
```



## ACLED

Extracted 2022-01-25. Conflict reference table, at incident level. Provided by MIMU. Read into .csv for public sharing. [Link](https://raw.githubusercontent.com/food-security-cluster-myanmar/exploratory-data-analysis-acled-fsc/master/data/acled_20220125.csv); only used for this report.  

Find updated dataset at [Link](). 

```{r acled}
acled <- read_excel("2010-01-01-2022-01-25-Myanmar.xlsx", 
         sheet = "2010-2021-Cleaned") %>% 
  mutate(admin3 = recode(admin3,
                        "Dagon Myothit-South" = "Dagon Myothit (South)",
                        "Dagon Myothit-North" = "Dagon Myothit (North)",
                        "Dagon Myothit-Seikkan" = "Dagon Myothit (Seikkan)",
                        "Dagon Myothit-East" = "Dagon Myothit (East)",
                        "Hlaingtharya-East" = "Hlaingtharya",
                        "Hlaingtharya-West" = "Hlaingtharya",
                        "Pangsang" = "Pangsang (Panghkam)")) %>% 
  # Locations in yangon with longlat, but no admin3 have been coded as Kyauktada, where city hall is;
  # I think you need to fix this. 
  # Similarly, in mandalay, they have been placed in Chanayethazan and Zay yar thi ri in NPT
  mutate(admin3 = case_when(is.na(admin3) & latitude == 16.8052 ~ "Kyauktada",
                           is.na(admin3) & latitude == 19.7450 ~ "Zay Yar Thi Ri",
                           is.na(admin3) & latitude == 21.9747 ~ "Chanayethazan",
                           TRUE ~ admin3)) %>% 
  left_join(townships %>%  
              select(township_name, admin3_pcode), by = c("admin3" = "township_name")) %>% 
  group_by(data_id) %>% 
  # The left_join somoetimes produces duplicates
  slice(which.min(event_id_no_cnty)) %>%
  ungroup() %>% 
  mutate(inter_type = interaction,
         inter_type = recode_factor(inter_type,
                             `10` = "SOLE	MILITARY ACTION",
                             `11` = "MILITARY VERSUS MILITARY",
                             `12` = "MILITARY VERSUS REBELS",
                             `13` = "MILITARY VERSUS POLITICAL MILITIA",
                             `14` = "MILITARY VERSUS COMMUNAL MILITIA",
                             `15` = "MILITARY VERSUS RIOTERS",
                             `16` = "MILITARY VERSUS PROTESTERS",
                             `17` = "MILITARY VERSUS CIVILIANS",
                             `18` = "MILITARY VERSUS OTHER",
                             `20` = "SOLE REBEL ACTION",
                             `22` = "REBELS VERSUS REBELS",
                             `23` = "REBELS	VERSUS POLITICAL MILIITA ",
                             `24` = "REBELS VERSUS COMMUNAL MILITIA",
                             `25` = "REBELS VERSUS RIOTERS",
                             `26` = "REBELS VERSUS PROTESTERS ",
                             `27` = "REBELS VERSUS CIVILIANS",
                             `28` = "REBELS VERSUS OTHERS",
                             `30` = "SOLE POLITICAL MILITIA ACTION",
                             `33` = "POLITICAL MILITIA VERSUS POLITICAL	MILITIA",
                             `34` = "POLITICAL MILITIA VERSUS COMMUNAL	MILITIA",
                             `35` = "POLITICAL MILITIA VERSUS RIOTERS",
                             `36` = "POLITICAL MILITIA VERSUS PROTESTERS",
                             `37` = "POLITICAL MILITIA VERSUS CIVILIANS",
                             `38` = "POLITICAL MILITIA VERSUS OTHERS",
                             `40` = "SOLE COMMUNAL MILITIA ACTION",
                             `44` = "COMMUNAL MILITIA VERSUS COMMUNAL MILITIA",
                             `45` = "COMMUNAL MILITIA VERSUS RIOTERS",
                             `46` = "COMMUNAL	MILITIA	VERSUS PROTESTERS ",
                             `47` = "COMMUNAL MILITIA VERSUS CIVILIANS",
                             `48` = "COMMUNAL MILITIA VERSUS OTHER",
                             `50` = "SOLE RIOTER ACTION",
                             `55` = "RIOTERS VERSUS RIOTERS",
                             `56` = "RIOTERS VERSUS PROTESTERS",
                             `57` = "RIOTERS VERSUS CIVILIANS",
                             `58` = "RIOTERS VERSUS OTHERS",
                             `60` = "SOLE PROTESTER ACTION",
                             `66` = "PROTESTERS VERSUS PROTESTERS",
                             `67` = "PROTESTERS VERSUS CIVILIANS",
                             `68` = "PROTESTERS VERSUS OTHER",
                             `78` = "OTHER ACTOR VERSUS CIVILIANS",
                             `80` = "SOLE OTHER ACTION"), 
         inter_type = tolower(str_trim(str_replace_all(inter_type, "[\\r\\n\\t]+", " ")))) %>% 
  mutate(admin3_pcode = ifelse(admin3_pcode == "MMR007014" & admin1 == "Yangon", "MMR013006", admin3_pcode),
         admin3_pcode = ifelse(admin1 == "Magway" & admin3 == "Minhla", "MMR009013", admin3_pcode)) %>% 
  mutate(has_fatalities = ifelse(fatalities > 0, TRUE, FALSE)) %>%
  mutate(month = lubridate::month(event_date, label = TRUE)) %>%
  mutate(non_combatant = case_when(event_type == "Violence against civilians" ~ TRUE, 
                                   str_detect(inter_type, "civilian|protester") ~ TRUE, 
                                   str_detect(notes, "civilian|protester|villager") ~ TRUE, 
                                   TRUE ~ FALSE))

acled %>% 
  write_csv("./data/acled_20220125.csv") 



```

## Actors 

Extracted 2022-02-09. Conflict reference table -- long table, by incident and actor. 
Only observations from 2021 used in the report. Public. [Report link](https://raw.githubusercontent.com/food-security-cluster-myanmar/exploratory-data-analysis-acled-fsc/master/data/actors.csv) (outdated); updated Link. 

```{r actors}
actors <- read_csv("1900-01-01-2022-02-09-Myanmar.csv") %>% 
  mutate(admin3 = recode(admin3,
                        "Dagon Myothit-South" = "Dagon Myothit (South)",
                        "Dagon Myothit-North" = "Dagon Myothit (North)",
                        "Dagon Myothit-Seikkan" = "Dagon Myothit (Seikkan)",
                        "Dagon Myothit-East" = "Dagon Myothit (East)",
                        "Hlaingtharya-East" = "Hlaingtharya",
                        "Hlaingtharya-West" = "Hlaingtharya",
                        "Pangsang" = "Pangsang (Panghkam)")) %>% 
  # locations in yangon with longlat, but no admin3 have been coded as Kyauktada, where city hall is;
  # similarly, in mandalay, they have been placed in Chanayethazan and Zay yar thi ri in NPT
  mutate(admin3 = case_when(is.na(admin3) & latitude == 16.8052 ~ "Kyauktada",
                           is.na(admin3) & latitude == 19.7450 ~ "Zay Yar Thi Ri",
                           is.na(admin3) & latitude == 21.9747 ~ "Chanayethazan",
                           TRUE ~ admin3)) %>% 
  left_join(townships %>%  
              select(township_name, admin3_pcode), by = c("admin3" = "township_name")) %>% 
  left_join(acled %>% 
              select(interaction, inter_type) %>%  distinct(), by = "interaction") %>% 
  mutate(actor_simple = actor1) %>% 
  mutate(actor_simple = 
           case_when(str_detect(actor_simple, "People's Defense Force") ~ "People's Defense Force",
                     str_detect(actor_simple, "Rohingya Muslim Group \\(Myanmar\\)") ~ "Rohingya Muslim Group (Myanmar)",
                     str_detect(actor_simple, "Women \\(Myanmar\\)") ~ "Women (Myanmar)",
                     str_detect(actor_simple, "21 Guerrilla Force") ~ "21 Guerrilla Force",
                     str_detect(actor_simple, "27 Revolution Force") ~ "27 Revolution Force",
                     str_detect(actor_simple, "88 Generation Peace and Open Society") ~ "88 Generation Peace and Open Society",
                     str_detect(actor_simple, "ABFSU") ~ "ABFSU",
                     str_detect(actor_simple, "Aid Workers \\(Myanmar\\)") ~ "Aid Workers (Myanmar)",
                     str_detect(actor_simple, "Ayadaw Bone Naing PDF") ~ "Ayadaw Bone Naing PDF",
                     str_detect(actor_simple, "Baptist Christian Group") ~ "Baptist Christian Group (Myanmar)",
                     str_detect(actor_simple, "Buddhist Group") ~ "Buddhist Group (Myanmar)", 
                     str_detect(actor_simple, "CDF: Chinland Defense Force") ~ "CDF: Chinland Defense Force",
                     str_detect(actor_simple, "Chin Ethnic Group") ~ "Chin Ethnic Group (Myanmar)",
                     str_detect(actor_simple, "Chinese Ethnic Group") ~ "Chinese Ethnic Group (Myanmar)",
                     str_detect(actor_simple, "Christian Group")~ "Christian Group (Myanmar)",
                     str_detect(actor_simple, "Civilians \\(Myanmar\\)") ~ "Civilians (Myanmar)",
                     str_detect(actor_simple, "CNO/CNDF") ~ "CNO/CNDF: Chin National Organization/Chin National Defense Force",
                     str_detect(actor_simple, "DKBA \\(Buddhist\\)") ~ "DKBA (Buddhist): Democratic Karen Buddhist Army",
                     str_detect(actor_simple, "Farmers \\(Myanmar\\)") ~ "Farmers (Myanmar)",
                     str_detect(actor_simple, "Democratic Guerrilla Force") ~ "Democratic Guerrilla Force",
                     str_detect(actor_simple, "DGWCF") ~ "DGWCF: Delta Guerrilla Warfare Collective Forces",
                     str_detect(actor_simple, "Former Government of Myanmar \\(2016-\\)") ~ "Former Government of Myanmar (2016-)",
                     str_detect(actor_simple, "Former Government of Myanmar \\(2021-\\)") ~ "Former Government of Myanmar (2021-)",
                     str_detect(actor_simple, "Former Military Forces of Myanmar \\(2016-2021\\)") ~ 
                       "Former Military Forces of Myanmar (2016-2021)",
                     str_detect(actor_simple, "Former Military Forces of Myanmar \\(2021-\\)") ~ 
                       "Former Military Forces of Myanmar (2021-)",
                     str_detect(actor_simple, "Government of Myanmar \\(2011-2016\\)") ~ "Government of Myanmar (2011-2016)",
                     str_detect(actor_simple, "Government of Myanmar \\(2016-\\)") ~ "Government of Myanmar (2016-)",
                     str_detect(actor_simple, "Government of Myanmar \\(2021-\\)") ~ "Government of Myanmar (2021-)",
                     str_detect(actor_simple, "Health Workers \\(Myanmar\\)") ~ "Health Workers (Myanmar)",
                     str_detect(actor_simple, "Journalists \\(Myanmar\\)") ~ "Journalists (Myanmar)",
                     str_detect(actor_simple, "Kachin Ethnic Group \\(Myanmar\\)") ~ "Kachin Ethnic Group (Myanmar)", 
                     str_detect(actor_simple, "Kayah Ethnic Group \\(Myanmar\\)")~ "Kayah Ethnic Group (Myanmar)",
                     str_detect(actor_simple, "Kayin Ethnic Group \\(Myanmar\\)")~ "Kayin Ethnic Group (Myanmar)",
                     str_detect(actor_simple, "Karenni Democratic Front") ~ "KDF: Karenni Democratic Front",
                     str_detect(actor_simple, "KIO/KIA") ~ "KIO/KIA: Kachin Independence Organization/Kachin Independence Army",
                     str_detect(actor_simple, "Karenni Nationalities Defense Force") ~ "KNDF: Karenni Nationalities Defense Force",
                     str_detect(actor_simple, "Karenni National Progressive Party") ~ "KNPP/KA: Karenni National Progressive Party/Karenni Army",
                     str_detect(actor_simple, "Karen National Union") ~ "KNU/KNLA: Karen National Union/Karen National Liberation Army",
                     str_detect(actor_simple, "Knights Of The Red Land") ~ "KORL: Knights Of The Red Land",
                     str_detect(actor_simple, "KNU/KNLA Peace Council") ~ "KPC: KNU/KNLA Peace Council",
                     str_detect(actor_simple, "LGBT \\(Myanmar\\)") ~ "LGBT (Myanmar)",
                     str_detect(actor_simple, "Labour Group \\(Myanmar\\)") ~ "Labour Group (Myanmar)",
                     str_detect(actor_simple, "Lisu Ethnic Group \\(Myanmar\\)") ~ "Lisu Ethnic Group (Myanmar)",
                     str_detect(actor_simple, "MG6 Guerrilla Force") ~ "MG6 Guerrilla Force",
                     str_detect(actor_simple, "MHG: Monywa Heroes Group") ~ "MHG: Monywa Heroes Group",
                     str_detect(actor_simple, "MHH: Mungshawa Hpyen Hpung") ~ "MHH: Mungshawa Hpyen Hpung",
                     str_detect(actor_simple, "Border Guard Force") ~ "Military Forces of Myanmar - Border Guard Force",
                     str_detect(actor_simple, "Police Forces of Myanmar") ~ "Police Forces of Myanmar",
                     str_detect(actor_simple, "People's Militia Force") ~ "People's Militia Force",
                     str_detect(actor_simple, "Military Forces of Myanmar") ~ "Military Forces of Myanmar",
                     str_detect(actor_simple, "MNTJP/MNDAA: Myanmar National Truth and Justice Party/Myanmar National Democratic Alliance Army") ~ 
                       "MNTJP/MNDAA: Myanmar National Truth and Justice Party/Myanmar National Democratic Alliance Army",
                     str_detect(actor_simple, "Mon Ethnic Group \\(Myanmar\\)") ~ "Mon Ethnic Group (Myanmar)",
                     str_detect(actor_simple, "Monywa Nightking") ~ "Monywa Nightking",
                     str_detect(actor_simple, "Monywa Young People Group") ~ "Monywa Young People Group",
                     str_detect(actor_simple, "Mro Ethnic Group \\(Myanmar\\)") ~ "Mro Ethnic Group (Myanmar)",
                     str_detect(actor_simple, "MTHSF: Mon Thaton Special Forces") ~ "MTHSF: Mon Thaton Special Forces",
                     str_detect(actor_simple, "Muslim Group \\(Myanmar\\)") ~ "Muslim Group (Myanmar)",
                     str_detect(actor_simple, "MWVA: Myanmar War Veterans' Association") ~ "MWVA: Myanmar War Veterans' Association",
                     str_detect(actor_simple, "NLD: National League for Democracy") ~ "NLD: National League for Democracy",
                     str_detect(actor_simple, "Pa-Oh Ethnic Group \\(Myanmar\\)") ~ "Pa-Oh Ethnic Group (Myanmar)",
                     str_detect(actor_simple, "Palaung/Ta'ang Ethnic Group \\(Myanmar\\)") ~ "Palaung/Ta'ang Ethnic Group (Myanmar)",
                     str_detect(actor_simple, "Pale Joker PDF") ~ "Pale Jokers PDF",
                     str_detect(actor_simple, "People's Defense Froce") ~ "People's Defense Force",
                     str_detect(actor_simple, "People Defense Froce") ~ "People's Defense Force",
                     str_detect(actor_simple, "People's Guerrillas Commando") ~ "People's Guerrillas Commando",
                     str_detect(actor_simple, "People's Militia Force") ~ "People's Militia Force",
                     str_detect(actor_simple, "Phoenix Defense Force") ~ "Phoenix Defense Force", 
                     str_detect(actor_simple, "Phoenix Special Guerrilla Force") ~ "Phoenix Special Guerrilla Force",
                     str_detect(actor_simple, "PNO/PNA: Pa-Oh National Organization/Pa-Oh National Army") ~ 
                       "PNO/PNA: Pa-Oh National Organization/Pa-Oh National Army",
                     str_detect(actor_simple, "Prisoners \\(Myanmar\\)") ~ "Prisoners (Myanmar)",
                     str_detect(actor_simple, "PSLF/TNLA: Palaung State Liberation Front/Ta'ang National Liberation Army") ~ 
                       "PSLF/TNLA: Palaung State Liberation Front/Ta'ang National Liberation Army",
                     str_detect(actor_simple, "Rakhine Ethnic Group \\(Myanmar\\)") ~ "Rakhine Ethnic Group (Myanmar)",
                     str_detect(actor_simple, "RCSS/SSA-S: Restoration Council of Shan State/Shan State Army-South") ~ 
                       "RCSS/SSA-S: Restoration Council of Shan State/Shan State Army-South",
                     str_detect(actor_simple, "Refugees/IDPs \\(Myanmar\\)") ~ "Refugees/IDPs (Myanmar)",
                     str_detect(actor_simple, "Shan Ethnic Group \\(Myanmar\\)") ~ "Shan Ethnic Group (Myanmar)",
                     str_detect(actor_simple, "Students \\(Myanmar\\)") ~ "Students (Myanmar)",
                     str_detect(actor_simple, "Teachers \\(Myanmar\\)") ~ "Teachers (Myanmar)",
                     str_detect(actor_simple, "ULA/AA: United League of Arakan/Arakan Army") ~ "ULA/AA: United League of Arakan/Arakan Army",
                     str_detect(actor_simple, "ULFA-I: United Liberation Front of Asom-Independent") ~ 
                       "ULFA-I: United Liberation Front of Asom-Independent",
                     str_detect(actor_simple, "Urban Guerrilla Group") ~ "Urban Guerrilla Group", 
                     str_detect(actor_simple, "USDP: Union Solidarity and Development Party") ~ "USDP: Union Solidarity and Development Party",
                     str_detect(actor_simple, "Zero Guerrilla Force") ~ "Zero Guerrilla Force",
                     str_detect(actor_simple, "DNP: Democratic Party of National Politics") ~ "DNP: Democratic Party of National Politics",
                     str_detect(actor_simple, "Hlaw Tet In Ahr Su 126: Rowing Force 126") ~ "Hlaw Tet In Ahr Su 126: Rowing Force 126",
                     TRUE ~ actor_simple)) %>%  
  mutate(inter1 = recode(inter1, 
                         `1` = "state_forces",
                         `2` = "rebel_groups",
                         `3` = "political_militias",
                         `4` = "identity_militias",
                         `5` = "rioters",
                         `6` = "protesters",
                         `7` = "civilians", 
                         `8` = "external_other_force")) %>%  
  mutate(inter1 = fct_relevel(inter1, c("external_other_force", "identity_militias", "rioters", "rebel_groups", 
                                        "political_militias", "state_forces", "civilians", "protesters"))) %>% 
  mutate(event_type = fct_relevel(event_type, c("Battles", "Explosions/Remote violence", "Violence against civilians",
                                                "Protests", "Riots", "Strategic developments"))) %>% 
  mutate(has_fatalities = ifelse(fatalities > 0, TRUE, FALSE))

actors %>%  
  write_csv("./data/actors.csv")
```

## ACLED Words

Textmining dataset. Created using `tidytext`. Contains all words from conflict descriptions of each conflict incident, filtered by stopword. Public. [Link](https://raw.githubusercontent.com/food-security-cluster-myanmar/exploratory-data-analysis-acled-fsc/master/data/acled_words.csv).

```{r}
# acled_words df
acled_words <- acled %>% 
  filter(year == 2021) %>% 
  select(data_id, notes) %>% 
  unnest_tokens(word, notes) %>% 
  anti_join(stop_words, by = "word") %>% 
  filter(str_detect(word, "[a-z]")) %>% 
  left_join(acled %>% 
              select(data_id, admin1, admin3_pcode, event_type, sub_event_type, fatalities, inter_type), by = "data_id")

acled_words %>%  
  write_csv("./data/acled_words.csv") 

```

## Vulnerability in Myanmar

Census reference dataset. Dataset used for the creation of the MIMU-HARP Vulnerability Index based on 2015 Census data. Public. [Link](https://raw.githubusercontent.com/food-security-cluster-myanmar/exploratory-data-analysis-acled-fsc/master/data/vulmmr.csv). 

```{r vulnerability in Myanmar}
# vulnerability in Myanmar 

vulmmr <- read_excel("Datasets_Vulnerability_Analysis_in_Myanmar_09Jul2018 (1).xlsx",
           skip = 1) %>% 
  slice(-c(1:3)) %>% 
  clean_names() %>% 
  select(-label) %>% 
  mutate_at(vars(number_of_village_tracts:wb_wealth_rank), as.numeric) %>% 
  mutate_at(vars(disasters_impacted_by_nargis_2008:acled_2015_2016_data_exists), as.logical) %>% 
  mutate_at(vars(conflict_2015_2016_number_of_battles:corrected_conflict_index_garry), as.numeric) %>% 
  select(-starts_with("x")) %>% 
  select(-c(private_sector_development_2014_2015, protection_2010_2015, shelter_2010_2015, wash_2010_2015))

vulmmr %>%  
  write_csv("./data/vulmmr.csv") 
```

